---
title: "Beer Data Analysis"
authors: "Daniel Davieau, Lakeitra Webb, Emil Ramos"
date: "February 26, 2018"
output:
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE)
```
## Introduction

The purpose of this documentaton is to provide a detailed breakdown related to several different Beers and the Breweries that produce them. The Beers dataset contains a list of 2410 US craft beers and the Breweries dataset contains 558 US breweries. The datasets descriptions are as follows:

The Environment Information section describes the  software environment and repository location of the data files and code. You should refer to this information should you desire to reproduce this analysis. 

Each subsequent section contains (in order) :  

* A particular question you have asked  

* Methods we used to analyze the data you provided  


* The answer to the question  

---  

## Environment Information  

>  This code block clears the R environment and loads the libraries required to process the subsequent code.  

```{r tidy=TRUE}
rm(list=ls())
library(ggplot2)
library(readr)
library(bitops)
library(plyr)
```
> This code block displays the hardware, software and thier versions.  

```{r tidy=TRUE}
sessionInfo()
```

[Link to the Github Repository Associated with this Study](https://github.com/davxdan/MSDS_6306_DoingDataScience_Case-Study_01)

--- 

### 1. How many breweries are present in each state?

#### Loading the Provided Data

```{r, tidy=TRUE}
#Load Beers.csv
RawBeerData <- read_csv("Input/RawDataFiles/Beers.csv")
```
```{r, tidy=TRUE}
#Load Breweries.csv
RawBreweryData <- read_csv("Input/RawDataFiles/Breweries.csv")
```
#### Examining the Provided Data
We examined the data using [OpenRefine](http://openrefine.org/) and other tools.  

We identified some records in the data files that raised concerns. For example the record (110,"Woodstock Inn, Station & Brewery",North Woodstock, NH) could potentially cause parsing errors however using the readr package mitigated this.

```{r, tidy=TRUE}
#Display the potentially troublesome records
RawBreweryData[c(110, 111, 112),]
#Load the raw data into a working dataframe. We want to preserve the raw data just in case we need it later.
Stage1BreweryData<- RawBreweryData 
#Convert the State column to character and isolate it so we can summarise it to get the counts by state. There are other ways to do this but this was simplest.
Stage1BreweryData <-transform(Stage1BreweryData, State = as.character(State)) 
CountBreweriesByState<-data.frame(Stage1BreweryData$State)
```
> Answer 1  

The count of Breweries by state are listed below:
```{r, tidy=TRUE}
summary(CountBreweriesByState,maxsum=100)
```
###2. Merge beer data with the breweries data. Print the ﬁrst 6 observations and the last six observations to check the merged ﬁle.

First, we needed to adjust the column names so we can successfully merge the brewery data with the beers data. The new column is titled, "Brewery_id."
```{r}
#Adjust the column names so we can merge brewery data with beers data on "Brewery_id".
colnames(Stage1BreweryData) <- c("Brewery_id","BreweryName","City","State") 
#Merge the data
Stage2<- merge(x= RawBeerData, y=Stage1BreweryData, by = c("Brewery_id"), all=FALSE) 
```
> Answer 2  

Below are the first and last six records of the resulting joined data:  
```{r, tidy=TRUE}
#display the first 6 rows
head(Stage2)
#display the last 6 rows
tail(Stage2)
```

###3. Report the number of NA’s in each column.

In order to summarize the amount of NA's in each column, we needed to create each variable. Once each were created, we were able to summarize a breakdown by column.
```{r, tidy=TRUE}
#Store the count of NA's in each data field in a variable
Brewery_id<-sum(is.na(Stage2$Brewery_id)) 
Name<-sum(is.na(Stage2$Name))
Beer_ID<-sum(is.na(Stage2$Beer_ID))
ABV<-sum(is.na(Stage2$ABV))
IBU<-sum(is.na(Stage2$IBU))
Style<-sum(is.na(Stage2$Style))
Ounces<-sum(is.na(Stage2$Ounces))
BreweryName<-sum(is.na(Stage2$BreweryName))
City<-sum(is.na(Stage2$City))
State<-sum(is.na(Stage2$State))
#Store the individual counts of NA's in a Matrix
NASummary<-as.matrix(c(Brewery_id, Name, Beer_ID, ABV, IBU, Style, Ounces, BreweryName, City, State))
#Name the column 
colnames(NASummary)<-c("Count of NA's")
#Name the rows
rownames(NASummary) <- c("Brewery_id", "Name", "Beer_ID", "ABV", "IBU", "Style", "Ounces", "BreweryName", "City", "State")
```
> Answer 3  

The count of NA's for each column are displayed below
```{r, tidy=TRUE}
#Display the data
NASummary
```

###4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r, tidy=TRUE}
#Create a function to get the medians
  getMedians <- function(x)
  {
  c(median = median(x,na.rm=TRUE ))
  }
#Create a dataframe and call the function to get ABV medians and store results in it
ABVMedians<-as.data.frame(tapply(Stage2$ABV, Stage2$State, getMedians))
#Create a dataframe and call the function to get IBU medians and store results in it
IBUMedians<-as.data.frame(tapply(Stage2$IBU, Stage2$State, getMedians))
```
>Answer 4  

The following graphs display the median alcohol content and international bitterness unit for each state for comparison.  

* Note: we excluded records that had no alchohol by volume and international bitterness unit values in the data provided

```{r, tidy=TRUE}
#Display the Median ABV by State 
par(las=2)
barplot(ABVMedians[,1],main ="Median Alchohol Content by State", horiz = FALSE, col = 4)
#Display the Median IBU by State 
par(las=2)
barplot(IBUMedians[,1],main ="International Bitterness Units by State", horiz = FALSE, col = 4)
```

###5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r, tidy=TRUE}
#get the maximum ABV by state. Excluding NA's
MaxABVState<-ddply(Stage2, .(State), summarise, MaxABVState = max(ABV, na.rm=TRUE)) 
#Sort by ABV
MaxABVState<-MaxABVState[order(MaxABVState$MaxABVState),]
#get the maximum IBU by state. Excluding NA's
MaxIBUState<-ddply(Stage2, .(State), summarise, MaxIBUState = max(as.double(Stage2$IBU), na.rm=TRUE))
#Sort by ABV
MaxIBUState<-MaxIBUState[order(MaxIBUState$MaxIBUState),]
```
>Answer 5  

The state with the highest alchohol by volume beer is Delaware with  $ABV = 0.055$  
The state with the highest international bitterness units is Arkansas with  $IBU = 138$  
* Note: we excluded records that had no alchohol by volume and international bitterness unit values in the data provided
```{r, tidy=TRUE}
head(MaxABVState,1)
head(MaxIBUState,1)
```
###6. Summary statistics for the ABV variable.

>Answer 6  

Below are summary statistics for alchohol by volume beers
```{r}
summary(Stage2$ABV)
```

###7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. You are welcome to use the ggplot2 library for graphs. Please ignore missing values in your analysis. Make your best judgment of a relationship and EXPLAIN your answer.

>Answer 7

The chart below indicates that there is possibly a relationship between the bitterness of the beer and its alcoholic content. We must emphasize the word possibly; we would need to know more about the samples before definitively stating there is a relationship. 

It seems possible (based on the data provided) that either: 

* bitterness influences alchohol content to be higher  

or  

* alchohol content influences bitterness to be higher.

```{r}
p <- ggplot(Stage2, aes(Stage2$ABV, Stage2$IBU))
p + geom_point(size = 1)
```

